\TheTitle

### 2019-12-13 Paper1/results.tex
updated pipe

For the last numerical study we again use the Euler equations, but this time not applied to the NACA airfoil, but rather to a bend Sod shock tube experiment. The prescribed initial condition describes a gas at rest, but with a discontinuity, i.e. as sudden jump, in density as well as energy in the upper part of the tube. The density $\rho_u = p/(R \cdot T)$ in the upper part results in $\rho_u = 1.289$ with $p = 101\;325$ Pa, a temperature of $273.15$ K and the specific gas constant for air $R=287.87$. The energy $\rho e_u$ is set to exactly $1.0$. For the initial conditions in the lower part we set $\rho_l = 0.5 \cdot \rho_u$ and $\rho e_l = 0.3$. The heat capacity ratio $\gamma$ is set to $1.4$ as in the previous studies. We again augment the deterministic case by inflicting three uncertainties this time. The applied boundary conditions are Euler slip conditions $\bm{v}^T\bm{n}=0$ for the walls of and Dirichlet conditions set to the initial conditions at each end of the tube. The first two uncertainties are the density and energy of the lower part of the shock tube. Here we set $\rho_l \sim U(1.189,1389)$ and $\rho e_l \sim U(0.2,0.4)$. The third uncertainty enters through the position of the shock itself and thus $y_\text{shock} \sim U(1.0, 1.2)$. Note, that in contrast to the upper testcases, here we are not interested in the steady state of the system but rather the evolution of the expectancy and variance in time. The computational mesh (see Figure \ref{fig:referenceSolutionsPipeMesh}) is an unstructured mesh of $25458$ triangular cells, where all the cells are similar in size. The simulations were run until a time of $0.2$s was exceeded and the corresponding results can be seen in Figure \ref{fig:solution3D}. The reference solution in Figure \ref{fig:referenceSolutionsPipe1} and \ref{fig:referenceSolutionsPipe2} was again computed using SC with 50 quadrature points in each stochastic dimension, yielding a total of $50^3=125\;000$ quadrature points.\\
For this testcase we want to compare the results for sparse grids w.r.t. tensorized grids in higher stochastical dimensions. As we made used of adaptivity, the corresponding moment orders and used quadrature points can best be seen in Table \ref{tab:orders3D}.

@@

	\item It was not possible to obtain a result for sparse grids with moments of order 3 or higher. Any level of the used sparse grids based on Clenshaw Curtis quadrature tested up to level 11 ($72\;705$ quadrature nodes) for order 3, resulted in a failure of the method while computing the Hessian during the Newton iteration.
	\item Sparse grids required a significant higher number of quadrature points for the sample order of moment in comparison to tensorized grids.
	\item When using sparse grids with adaptivity the computational cost per quadrature node was observed to be smaller, proving that the nestedness property can potentially increase the methods performance.

@@

Even though the solutions in Figure \ref{fig:solution3D} show the same characteristics and sparse grids are generally well suited and often used in combination with the SC method, our tests reveal that, at least for the Clenshaw-Curtis based sparse grids we have used, these quadrature rules are not well suited to be used in combination with IPM as they require significantly more quadrature points and only give results for low orders of moments. However, as three stochastical dimensions are difficult to deal with in hyperbolical problems, but not for an UQ analysis in general, this might change for high dimensions.

>>>>

For the last numerical study we again use the Euler equations, but this time not in the context of the NACA airfoil, but rather with a bend Sod shock tube experiment. The prescribed initial condition for this testcase describes a gas at rest, but with a discontinuity, in density as well as energy in the upper part of the tube. The density $\rho_u = p/(R \cdot T)$ in the upper part is set to $\rho_u = 1.289$ with $p = 101\;325$ Pa, a temperature of $273.15$ K and the specific gas constant for dry air $R=287.87$. The energy $\rho e_u$ is set to exactly $1.0$. For the initial conditions in the lower part we set $\rho_l = 0.5 \cdot \rho_u$ and $\rho e_l = 0.3$. The heat capacity ratio $\gamma$ is set to $1.4$ as in the previous studies. We again augment the deterministic case by inflicting uncertainties and again increase the number uncertainties to three. The applied boundary conditions are Euler slip conditions $\bm{v}^T\bm{n}=0$ for the walls and Dirichlet type boundary conditions at the ends of the tube which set to the deterministic initial conditions. The first two uncertainties are the density and energy of the lower part of the shock tube. Here we set $\rho_l \sim U(1.189,1389)$ and $\rho e_l \sim U(0.2,0.4)$. The third uncertainty enters through the position of the shock itself and thus $y_\text{shock} \sim U(1.0, 1.2)$. Note, that in contrast to the upper testcases, here we are not interested in the steady state of the system, but rather the of time evolution of the expectancy and variance. The computational mesh (see Figure \ref{fig:referenceSolutionsPipeMesh}) is an unstructured mesh of $25\;458$ triangular cells, where the cells are all similar in size, i.e. there are not refined regions as in the NACA testcase. The simulations were run until a time of $0.2$s. The corresponding results can be seen in Figure \ref{fig:solution3D}. The reference solution in Figure \ref{fig:referenceSolutionsPipe1} and \ref{fig:referenceSolutionsPipe2} was computed using SC with 50 quadrature points in each stochastic dimension, yielding a total of $50^3=125\;000$ quadrature points.\\
For this testcase we want to compare the results for sparse grids w.r.t. tensorized grids in higher stochastical dimensions. As we made used of adaptivity, the corresponding moment orders and used quadrature points can be seen in Table \ref{tab:orders3D}.

@@

	\item It was not possible to obtain a result for sparse grids with moments of order 3 or higher. Any level up to level 11 ($72\;705$ quadrature nodes) of the used Clenshaw-Curtis sparse grids, resulted in a failure of the method during the computation of the mapping from the dual variables to the moment vector.
	\item Sparse grids required a significantly higher number of quadrature points for the same order of moments in comparison to tensorized grids.
	\item When using sparse grids with adaptivity the computational cost per quadrature node was observed to be smaller, showing that the nestedness property can potentially increase the methods performance.

@@

Even though the solutions in Figure \ref{fig:solution3D} show the same characteristics and sparse grids are generally well suited and often used in combination with the SC method, our tests reveal that, at least for the Clenshaw-Curtis based sparse grids, these quadrature rules are not well suited to be used in combination with IPM as they require significantly more quadrature points and are only able to generate results for low orders of moments. Tensorized quadrature on the other hand performs similarly well as in the previous testcases and should thus be preferred for lower dimensions. 
